Airspace configuration decision based on Monte-Carlo tree search

### 2018-04-16 article/article.tex
(+) Added par on selection and expansion

\subsection{Monte Carlo tree search}
\subsubsection{Overview}
Monte-Carlo is a best-first search method using stochastic simulations. The
algorithm actually uses two trees, an underlying tree associated to the model
(e.g.\ a game tree) and a search tree. The latter is built incrementally by the
algorithm the following way, at each step, starting from the root, the most
promising nodes towards the bottom of the tree are selected. Once a leaf of the
search tree is reached, a new node coming from the model tree is added to the
leaf. A simulation is then run from this newly added node to evaluate the
outcome of a path starting from it. The result of the simulation is eventually
backpropagated to the consecutive fathers of the new node.
\begin{compactenum}
    \item selection,
    \item expansion,
    \item simulation,
    \item backpropagation.
\end{compactenum}
Those phases are repeated until a stopping criterium is reached,
see\ref{alg:gen_mcts} (memory usage or time).
\begin{algorithm}
    \caption{General MCTS~\cite{browne2012survey}}\label{alg:gen_mcts}
    \begin{algorithmic}
        \Procedure{MctsSearchTree}{$v_0$}
        \While{within computational budget}
        \State{} \([v^l_1, \dots, v^l_i] \gets\) \Call{TreePolicy}{$v_0$}
        \State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v^l_i$}
        \State{} \(r \gets\) \Call{Reward}{$\Delta$}
        \State{} \Call{Backup}{$[v^l_1, \dots, v^l_i]$, $r$}
        \EndWhile{}
        \EndProcedure{}
    \end{algorithmic}
\end{algorithm}

\subsection{UCT selection}
\begin{algorithm}
    \caption{Extending the search tree}
    \begin{algorithmic}
        \Function{TreePolicy}{$v$}
        \State{} \([v_1, \dots, v_n] \gets\) \Call{Select}{$v$}
        \If{$v_n$ is terminal}
        \State{} \Return{$[v_1, \dots, v_n]$}
        \Else{}
        \State{} \([v_1, \dots, v_{n+1}] \gets\)
        \Call{Expand}{$[v_1, \dots, v_n]$}
        \State{} \Return{$[v_1, \dots, v_{n+1}]$}
        \EndIf{}
        \EndFunction{}
    \end{algorithmic}
    \begin{algorithmic}
        \Function{Select}{$v, a$}
        \If{$v$ is terminal or fully expanded}
        \State{} \Return{$[v, a]$}
        \Else{}
        \State{} fav\(\gets\) \Call{BestChild}{$v$}
        \State{} \Return{\Call{Select}{fav, $[v, a]$}}
        \EndIf{}
        \EndFunction{}
    \end{algorithmic}
    \begin{algorithmic}
        \Function{BestChild}{$v$}
        \State{} \Return{$\argmax\{\Call{UCB}{v'} |
        v' \text{\,children of\,} v\}$}
        \EndFunction{}
    \end{algorithmic}
\end{algorithm}

@@

In this section, we discuss the model established to approach this configuration
problem. First we defined a structure for a controled sector and all possible
transitions through time. Then a workload model is defined. And finally we
determined a cost per partition and a cost function aimed to be minimized.

\subsection{State}

\subsection{workload}

$p_{low}$
$p_{normal}$
$p_{high}$

\subsection{Partition cost and cost function}

In order to evaluate a given partition $P$ at time $t$, a cost $C(P, t)$ needs
to be defined. This cost depend on the workload in each the sector. We give the
following intermediate definitions to represent a high (respectively normal
and low) cost for the partition :
\begin{itemize}
    \item $c_+(P, t) = \sum_{S \in P_t} \delta_{h}(S, t) * p_{high}^{S, t} * |S|^2$
    \item $c_=(P, t) = \sum_{S \in P_t} \delta_{n}(S, t) * p_{norm}^{S} * |S|^{-2}$
    \item $c_-(P, t) = \sum_{S \in P_t} \delta_{l}(S, t) * p_{low}^{S, t} * |S|^{-2}$
\end{itemize}
with $\delta_{h}(P, t)$ (resp. $\delta_{n}(P_t)$ and $\delta_{l}(P_t)$) equal 1
if the probability $p_{high}$ (resp. $p_{normal}$ and $p_{low}$) is superior to
the two other, and 0 otherwise.

With

\section*{Notations}

\begin{itemize}
    \item $m$ : elementary module
    \item $S$ : sector control position, group of modules
    \item $P$ : partition of the airspace, group of sectors
    \item $t$ : current time
    \item $C(P, t)$ : Cost of a given partition $P$ at time $t$
    \item $s$ : state, in this article a partition $P$ and a time $t$
    \item $h(s)$ : heuristic that gives for a state $s$ an estimation of the cost
    \item $\mu_{F, a}$ : mean of all reward comming from node $F$ with action $a$
    \item $T_F$ : number of simuation where the final path include the node $F$
    \item $T_{F, a}$ : number of simuation where the final path include the node $F$ with action $a$
\end{itemize}

>>>>

\subsection{Monte Carlo tree search}
\subsubsection{Overview}
Monte-Carlo is a best-first search method using stochastic simulations. The
algorithm actually uses two trees, an underlying tree associated to the model
(e.g.\ a game tree) and a search tree. The latter is built incrementally by the
algorithm the following way, at each step, starting from the root, the most
promising nodes towards the bottom of the tree are selected. Once a leaf of the
search tree is reached, a new node coming from the model tree is added to the
leaf. A simulation is then run from this newly added node to evaluate the
outcome of a path starting from it. The result of the simulation is eventually
backpropagated to the consecutive fathers of the new node.

The above mentioned steps are denoted
\begin{compactenum}
	\item selection (or tree walk): choosing successively most promising nodes,
	\item expansion: adding new nodes to the search tree,
	\item simulation (or random walk): choosing quickly actions from the model
		tree,
	\item backpropagation (or backup): applying the result of the simulation to
		the search tree.
\end{compactenum}
Those phases are repeated until a stopping criterium (e.g.\ memory or time) is
reached, resulting in algorithm~\ref{alg:gen_mcts}.
\begin{algorithm}
	\caption{General MCTS~\cite{browne2012survey}}\label{alg:gen_mcts}
	\begin{algorithmic}
		\Procedure{MctsSearchTree}{$v_0$}
		\While{within computational budget}
		\State{} \([v_0, \dots, v_n] \gets\) \Call{TreePolicy}{$v_0$}
		\State{} \(\Delta \gets\) \Call{DefaultPolicy}{$v_n$}
		\State{} \Call{Backup}{$[v_0, \dots, v_n]$, $\Delta$}
		\EndWhile{}
		\EndProcedure{}
	\end{algorithmic}
\end{algorithm}

\subsubsection{Extending the search tree}
\paragraph{Selection}
The aim of the selection is to choose from a set of actions, the most promising
one\footnote{i.e.\ possibly leading to the best evaluation} according to a
\emph{tree policy}. This type of problem is addressed by bandits methods where
a multi armed bandit aims to always activate the slot machine giving the highest
reward.

The bandit problem is applied to MCTS via the Upper Confidence Tree algorithm.
It uses the UCB for Upper Confidence Bound as tree policy (introduced
in~\cite{kocsis2006bandit}). Let \(F\) be the node from which an action must be
selected to go deeper in the tree, \(a\) an action available from \(F\),
\(\mu_{F,a}\) the expected outcome from the node obtained by applying action
\(a\) on \(F\), \(T_F\) the number of simulations carried out from node \(F\)
(which includes any simulation from nodes in any subtree of \(F\)) and
\(T_{F,a}\) the number of simulations carried out from the node resulting from
the action \(a\) applied to node \(F\), \(\beta\) a constant. The selected node
is thus the one maximising
\begin{equation}
	\text{UCB} = \mu_{F,a} + 2 \beta \sqrt{\frac{2 \log T_f}{T_{F,a}}}
\end{equation}

\paragraph{Expansion}
The above step is carried out only if, considering a node \(F\), a simulation
has been run through each child available from \(F\). In the case where one or
more children haven't been explored yet, one of the unexplored children is
chosen and a simulation is started from it.
\begin{algorithm}
	\caption{Extending the search tree}
	\begin{algorithmic}
		\Function{TreePolicy}{$v$}
		\State{} \([v_1, \dots, v_n] \gets\) \Call{Select}{$v$}
		\If{$v_n$ is terminal}
		\State{} \Return{$[v_1, \dots, v_n]$}
		\Else{}
		\State{} \([v_1, \dots, v_{n+1}] \gets\)
		\Call{Expand}{$[v_1, \dots, v_n]$}
		\State{} \Return{$[v_1, \dots, v_{n+1}]$}
		\EndIf{}
		\EndFunction{}
	\end{algorithmic}
	\begin{algorithmic}
		\Function{Select}{$v, a$}
		\If{$v$ is terminal or fully expanded}
		\State{} \Return{$[v, a]$}
		\Else{}
		\State{} fav\(\gets\) \Call{BestChild}{$v$}
		\State{} \Return{\Call{Select}{fav, $[v, a]$}}
		\EndIf{}
		\EndFunction{}
	\end{algorithmic}
	\begin{algorithmic}
		\Function{BestChild}{$v$}
		\State{} \Return{$\argmax\{\Call{UCB}{v'} |
		v' \text{\,children of\,} v\}$}
		\EndFunction{}
	\end{algorithmic}
\end{algorithm}

@@
